STR: Seasonal-Trend decomposition using Regression

### 2020-11-17 str_submission.tex
Submitted version

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={STR: Seasonal-Trend decomposition using Regression},
            pdfkeywords={time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{strrefs.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{STR: Seasonal-Trend decomposition using Regression}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage{mathpazo}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \placefig{2}{1.5}{width=5cm}{monash2}
        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
        \begin{textblock}{7}(12.7,27.9)\hfill
        \includegraphics[height=0.7cm]{AACSB}~~~
        \includegraphics[height=0.7cm]{EQUIS}~~~
        \includegraphics[height=0.7cm]{AMBA}
        \end{textblock}
        \vspace*{2cm}
        \begin{center}\Large
        Department of Econometrics and Business Statistics\\[.5cm]
        \footnotesize http://monash.edu/business/ebs/research/publications
        \end{center}\vspace{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
\let\maketitle\titlepage
\makeatother

%% SPACING
\RequirePackage{setspace}
\spacing{1.5}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

%% KEYWORDS
\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

%% ABSTRACT
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
  {\end{minipage}}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

\nojel

\RequirePackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}

\nocover

\author{Alexander~Dokumentov, Rob J~Hyndman}
\addresses{\textbf{Alexander Dokumentov}\newline
Growing Data\newline Australia
\\[1cm]
\textbf{Rob J Hyndman}\newline
Monash University\newline Australia
\newline{Email: \href{mailto:Rob.Hyndman@monash.edu}{\nolinkurl{Rob.Hyndman@monash.edu}}}\newline Corresponding author\\[1cm]
}

\date{\sf\Date~\Month~\Year}
\makeatletter
 \lfoot{\sf Dokumentov, Hyndman: \@date}
\makeatother

\graphicspath{{figs/}}
\def\addlinespace{}
\usepackage{amsmath}
\setlength{\arraycolsep}{3pt}
\setcounter{MaxMatrixCols}{20}
\newcommand{\argmin}{\operatorname{arg\,min}}
\newcommand{\sn}{\kappa}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\diag}{\operatorname{diag}}
\renewcommand{\mod}{\operatorname{mod}}
\usepackage[section]{placeins}

\begin{document}
\maketitle
\begin{abstract}
We propose a new method for decomposing seasonal data: STR (a Seasonal-Trend decomposition using Regression). Unlike other decomposition methods, STR allows for multiple seasonal and cyclic components, covariates, seasonal patterns that may have non-integer periods, and seasonality with complex topology. It can be used for time series with any regular time index including hourly, daily, weekly, monthly or quarterly data. It is competitive with existing methods when they exist, but tackles many more decomposition problem than other methods allow.

\par

STR is based on a regularized optimization, and so is somewhat related to ridge regression. Because it is based on a statistical model, we can easily compute confidence intervals for components, something that is not possible with most existing decomposition methods (such as STL, X-12-ARIMA, SEATS-TRAMO, etc.).

\par

Our model is implemented in the R package \emph{stR}, so can be applied by anyone to their own data.
\end{abstract}
\begin{keywords}
time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM
\end{keywords}

\hypertarget{time-series-decomposition}{%
\section{Time series decomposition}\label{time-series-decomposition}}

@@

Existing time series decomposition methods are designed for monthly and quarterly data, with few tools available for more frequent data. The STL procedure of \textcite{cleveland1990stl} is the only widely available decomposition tool for data observed more frequently than monthly, but it assumes a very simple additive structure and allows no covariates. This makes it unsuitable for a large number of problems.

@@

The first attempts to decompose time series into various components can be dated to the 19th century, when \textcite{Poynting1884} proposed price averaging as a tool for eliminating trend and seasonal fluctuations. Later, his approach was extended by \textcite{hooker1901suspension}, \textcite{spencer1904graduation} and \textcite{anderson1914nochmals}. Further research in this direction included that by \textcite{copeland1915statistical}, who was the first to attempt to extract the seasonal component, until \textcite{macaulay1930smooth} proposed a method which gradually became ``classical''. The work of Macaulay led to the Census II method \autocite{shiskin1957electronic} which became popular amongst national statistics offices. The Census II method has continued to evolve, with additional features including robustness, calendar effects, covariates, ARIMA extensions, and extensive diagnostics. Widely used versions of this approach have been X-11 \autocite{shishkin1967x}, X-11-ARIMA \autocite{dagum1988x11arima,ladiray2001seasonal}, X-12-ARIMA \autocite{findley1998new} and X-13-ARIMA-SEATS \autocite{findley2005some}.

A different approach was followed by \textcite{cleveland1990stl}, who developed STL (Seasonal-Trend decomposition using Loess), which has come to be widely used outside the national statistics agencies, partly because of its availability in R \autocite{R}. This method uses iterative Loess smoothing to obtain an estimate of the trend, and then Loess smoothing again to extract a changing additive seasonal component.

Several model-based methods for seasonal decomposition have been developed, including the TRAMO/SEATS procedure which was developed at the Bank of Spain \autocite{Gomez2001}, the TBATS model of \textcite{de2011forecasting}, and various structural time series model approaches \autocite{harvey1990forecasting,Commandeur2011}. One big advantage of using a model for seasonal decomposition and adjustment is that it provides a natural way to compute confidence and prediction intervals.

@@

Where components behave multiplicatively, we can use a log or Box-Cox transformation \autocite{BC64} to made the additive assumption more reasonable.

@@

Smoothness of a function implies that its second derivatives are small, since otherwise large values of the second derivatives would lead to high curvature of the function making it ``wiggly'' \autocite[see][]{Wood}. Thus, we consider a function smooth if its second derivatives are normally distributed with zero mean and where the variance controls the degree of smoothness.

@@

Equivalently, the topology shown in Figure \ref{fig:topology2_pdf} can be described as two connected cylinders.

@@

Covariates are often available to help explain some of the variation seen in a time series. For example, \textcite{findley2009stock} considers the effects of various moving holidays on human activities, and \textcite{bell2004modeling} considers covariates with time-varying coefficients.

@@

We will use cross-validation for estimating the \(\lambda\) and \(\theta\) parameters of \eqref{regr:X}. Since the model in \eqref{STR:linear_model} is a linear model, the leave-one-out cross-validation residuals can be calculated \autocite[see][]{seber2012linear} using
\begin{equation}
  \label{FindLambdas:residuals}
  \operatorname{e}_{(i)} = \frac{y_i-\hat{y_i}}{1-h_{ii}}\ ,
\end{equation}
where \(y_i\) is the \(i\)th element of the vector \(\bm{y}\), \(\hat{\bm y}_i\) is the \(i\)th element of the vector \(\hat{\bm y} = \bm{H}\bm{y}\), and \(h_{ii}\) is the \(i\)th diagonal element of the hat matrix \(\bm{H} = \bm{X}(\bm{X}'\bm{X})^{-1}\bm{X}'\).
Therefore, we can use the well-known formula for cross-validation in linear regression \autocite[see, for example,][p.45]{RWC02}:
\begin{equation}
  \label{FindLambdas:CV}
  \text{CV} = \sum_{i=1}^{n}\left(\frac{y_i-\hat{y_i}}{1-h_{ii}}\right)^2.
\end{equation}

STR finds the optimal \(\lambda\) and \(\theta\) smoothing parameters by minimising CV. The problem of minimising CV can be complex, for example if there are many local minima, and we have no method that guarantees the finding of the global minimum. However, rather simple methods often work well in practice. We perform such optimisation using the Nelder-Mead method as implemented in the \texttt{optim()} function from the stats package in R \autocite{R}.

@@

\begin{figure}
\centering
\includegraphics{str_ijds_files/figure-latex/supermarket-1.pdf}
\caption{\label{fig:supermarket}Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\% confidence intervals for the trend and season components in blue}
\end{figure}

@@

We obtained similar results using other decomposition methods applied to the same data including STL \autocite{cleveland1990stl} and X-13-ARIMA-SEATS \autocite{findley2005some}.

\FloatBarrier

@@

This data set has two seasonal patterns. The first pattern is a weekly seasonal pattern (\(m_1=48\times7=336\)) that represents the specific demand features that are attributable to a particular day of the week. The second pattern is a daily seasonal pattern (\(m_2=48\)) with the topology in Figure \ref{fig:topology2_pdf}, which allows the model to distinguish between working days and holidays/weekends, and to make transitions between them. The pattern reflects the tendency to have a higher electricity demand during standard working hours and a lower demand at night. It also reflects the tendency to have different demand patterns on working days and holidays/weekends. A longer series would also have an annual seasonal pattern, but with only 115 days, this cannot be distinguished from the trend component.

\begin{figure}
\centering
\includegraphics{str_ijds_files/figure-latex/vic-1.pdf}
\caption{\label{fig:vic}Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.}
\end{figure}

@@

\begin{table}

\caption{\label{tab:outlierstable}The ten residuals that are largest in absolute value after an STR decomposition.}
\centering
\begin{tabular}[t]{rlcr}
\toprule
Date & Day of week & Time period & Residual\\
\midrule
3 February 2000 & Thursday & 14:30 -- 15:00 & 687.3\\
3 February 2000 & Thursday & 15:00 -- 15:30 & 871.3\\
3 February 2000 & Thursday & 15:30 -- 16:00 & 937.0\\
3 February 2000 & Thursday & 16:00 -- 16:30 & 767.2\\
3 February 2000 & Thursday & 16:30 -- 17:00 & 580.4\\

4 February 2000 & Friday & 16:00 -- 16:30 & -569.7\\
10 February 2000 & Thursday & 14:30 -- 15:00 & -611.4\\
10 February 2000 & Thursday & 15:00 -- 15:30 & -658.9\\
10 February 2000 & Thursday & 15:30 -- 16:00 & -695.7\\
10 February 2000 & Thursday & 16:00 -- 16:30 & -728.5\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{str_ijds_files/figure-latex/MelTemp-1.pdf}
\caption{\label{fig:MelTemp}The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day.}
\end{figure}

@@

Similarly, seasonal functional components can be obtained, using Fourier terms for example \autocite[reminiscent of][]{de2011forecasting}. This approach leads, with no additional effort, to handling seasonality with a fractional or varying period.

@@

This can be written as a quantile regression \autocite{koenker2005quantile}:
\begin{equation}
  \label{RSTR:qr}
  \hat{\bm\eta} = \operatorname{arg\,min}\limits_{\bm\eta} \left\|\bm{y}^{+}-\bm{X}\bm{\eta}\right\|_{L_1}\ ,
\end{equation}
where \(\bm{y}^{+}\), \(\bm{X}\) and \(\bm\eta\) are defined as in \eqref{STR:linear_model}. This can be solved numerically using a quantile regression algorithm \autocite{quantregR}.

>>>>

\documentclass[ijds,nonblindrev]{informs-ijds}

\OneAndAHalfSpacedXI
%%\OneAndAHalfSpacedXII % Current default line spacing
%%\DoubleSpacedXII
%%\DoubleSpacedXI

% Private macros here (check that there is no clash with the style)

\graphicspath{{figs/}}
\def\addlinespace{}
\usepackage{amsmath,bm,hyperref,url,booktabs}
\setlength{\arraycolsep}{3pt}
\setcounter{MaxMatrixCols}{20}
\renewcommand{\vec}{\operatorname{vec}}
\newcommand{\diag}{\operatorname{diag}}
\renewcommand{\mod}{\operatorname{mod}}
\usepackage[section]{placeins}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Natbib setup for author-year style
\usepackage{natbib}
 \bibpunct[, ]{(}{)}{,}{a}{}{,}%
 \def\bibfont{\small}%
 \def\bibsep{\smallskipamount}%
 \def\bibhang{24pt}%
 \def\newblock{\ }%
 \def\BIBand{and}%

%% Setup of theorem styles. Outcomment only one.
%% Preferred default is the first option.
\TheoremsNumberedThrough     % Preferred (Theorem 1, Lemma 1, Theorem 2)
%\TheoremsNumberedByChapter  % (Theorem 1.1, Lema 1.1, Theorem 1.2)
\ECRepeatTheorems

%% Setup of the equation numbering system. Outcomment only one.
%% Preferred default is the first option.
\EquationsNumberedThrough    % Default: (1), (2), ...
%\EquationsNumberedBySection % (1.1), (1.2), ...

% For new submissions, leave this number blank.
% For revisions, input the manuscript number assigned by the on-line
% system along with a suffix ".Rx" where x is the revision number.
\MANUSCRIPTNO{}

\begin{document}

% Outcomment only when entries are known. Otherwise leave as is and
%   default values will be used.
%\setcounter{page}{1}
%\VOLUME{00}%
%\NO{0}%
%\MONTH{Xxxxx}% (month or a similar seasonal id)
%\YEAR{0000}% e.g., 2005
%\FIRSTPAGE{000}%
%\LASTPAGE{000}%
%\SHORTYEAR{00}% shortened year (two-digit)
%\ISSUE{0000} %
%\LONGFIRSTPAGE{0001} %
%\DOI{10.1287/xxxx.0000.0000}%

% Author's names for the running heads
% Sample depending on the number of authors;
% \RUNAUTHOR{Jones}
% \RUNAUTHOR{Jones and Wilson}
% \RUNAUTHOR{Jones, Miller, and Wilson}
% \RUNAUTHOR{Jones et al.} % for four or more authors
% Enter authors following the given pattern:
%\RUNAUTHOR{}

% Title or shortened title suitable for running heads. Sample:
% \RUNTITLE{Bundling Information Goods of Decreasing Value}
% Enter the (shortened) title:
\RUNTITLE{STR: Seasonal-Trend decomposition using Regression}

% Full title. Sample:
% \TITLE{Bundling Information Goods of Decreasing Value}
% Enter the full title:
\TITLE{STR: Seasonal-Trend decomposition using Regression}

% Block of authors and their affiliations starts here:
% NOTE: Authors with same affiliation, if the order of authors allows,
%   should be entered in ONE field, separated by a comma.
%   \EMAIL field can be repeated if more than one author
\ARTICLEAUTHORS{%
\AUTHOR{Alexander Dokumentov}
\AFF{Growing Data, Australia} %, \URL{}}
\AUTHOR{Rob J Hyndman}
\AFF{Department of Econometrics \& Business Statistics, Monash University, Clayton VIC 3800, Australia, \EMAIL{Rob.Hyndman@monash.edu}}
% Enter all authors
} % end of the block

\ABSTRACT{%
We propose a new method for decomposing seasonal data: STR (a Seasonal-Trend decomposition using Regression). Unlike other decomposition methods, STR allows for multiple seasonal and cyclic components, covariates, seasonal patterns that may have non-integer periods, and seasonality with complex topology. It can be used for time series with any regular time index including hourly, daily, weekly, monthly or quarterly data. It is competitive with existing methods when they exist, but tackles many more decomposition problem than other methods allow.\par
STR is based on a regularized optimization, and so is somewhat related to ridge regression. Because it is based on a statistical model, we can easily compute confidence intervals for components, something that is not possible with most existing decomposition methods (such as STL, X-12-ARIMA, SEATS-TRAMO, etc.).\par
Our model is implemented in the R package \emph{stR}, so can be applied by anyone to their own data.
}%

% Sample
%\KEYWORDS{deterministic inventory theory; infinite linear programming duality;
%  existence of optimal policies; semi-Markov decision process; cyclic schedule}

% Fill in data. If unknown, outcomment the field
\KEYWORDS{time series decomposition, seasonal data, Tikhonov regularisation, ridge regression, LASSO, STL, TBATS, X-12-ARIMA, BSM}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Samples of sectioning (and labeling) in MNSC
% NOTE: (1) \section and \subsection do NOT end with a period
%       (2) \subsubsection and lower need end punctuation
%       (3) capitalization is as shown (title style).
%
%\section{Introduction.}\label{intro} %%1.
%\subsection{Duality and the Classical EOQ Problem.}\label{class-EOQ} %% 1.1.
%\subsection{Outline.}\label{outline1} %% 1.2.
%\subsubsection{Cyclic Schedules for the General Deterministic SMDP.}
%  \label{cyclic-schedules} %% 1.2.1
%\section{Problem Description.}\label{problemdescription} %% 2.

% Text of your paper here

\section{Time series decomposition}\label{time-series-decomposition}

@@

Existing time series decomposition methods are designed for monthly and quarterly data, with few tools available for more frequent data. The STL procedure of \citet{cleveland1990stl} is the only widely available decomposition tool for data observed more frequently than monthly, but it assumes a very simple additive structure and allows no covariates. This makes it unsuitable for a large number of problems.

@@

The first attempts to decompose time series into various components can be dated to the 19th century, when \citet{Poynting1884} proposed price averaging as a tool for eliminating trend and seasonal fluctuations. Later, his approach was extended by \citet{hooker1901suspension}, \citet{spencer1904graduation} and \citet{anderson1914nochmals}. Further research in this direction included that by \citet{copeland1915statistical}, who was the first to attempt to extract the seasonal component, until \citet{macaulay1930smooth} proposed a method which gradually became ``classical''. The work of Macaulay led to the Census II method \citep{shiskin1957electronic} which became popular amongst national statistics offices. The Census II method has continued to evolve, with additional features including robustness, calendar effects, covariates, ARIMA extensions, and extensive diagnostics. Widely used versions of this approach have been X-11 \citep{shishkin1967x}, X-11-ARIMA \citep{dagum1988x11arima,ladiray2001seasonal}, X-12-ARIMA \citep{findley1998new} and X-13-ARIMA-SEATS \citep{findley2005some}.

A different approach was followed by \citet{cleveland1990stl}, who developed STL (Seasonal-Trend decomposition using Loess), which has come to be widely used outside the national statistics agencies, partly because of its availability in R \citep{R}. This method uses iterative Loess smoothing to obtain an estimate of the trend, and then Loess smoothing again to extract a changing additive seasonal component.

Several model-based methods for seasonal decomposition have been developed, including the TRAMO/SEATS procedure which was developed at the Bank of Spain \citep{Gomez2001}, the TBATS model of \citet{de2011forecasting}, and various structural time series model approaches \citep{harvey1990forecasting,Commandeur2011}. One big advantage of using a model for seasonal decomposition and adjustment is that it provides a natural way to compute confidence and prediction intervals.

@@

Where components behave multiplicatively, we can use a log or Box-Cox transformation \citep{BC64} to made the additive assumption more reasonable.

@@

Smoothness of a function implies that its second derivatives are small, since otherwise large values of the second derivatives would lead to high curvature of the function making it ``wiggly'' \citep[see][]{Wood}. Thus, we consider a function smooth if its second derivatives are normally distributed with zero mean and where the variance controls the degree of smoothness.

@@

Equivalently, the topology shown in Figure \ref{fig:topology_pdf} can be described as two connected cylinders.

@@

Covariates are often available to help explain some of the variation seen in a time series. For example, \citet{findley2009stock} considers the effects of various moving holidays on human activities, and \citet{bell2004modeling} considers covariates with time-varying coefficients.

@@

We will use cross-validation for estimating the \(\lambda\) and \(\theta\) parameters of \eqref{regr:X}. Since the model in \eqref{STR:linear_model} is a linear model, the leave-one-out cross-validation residuals can be calculated \citep[see][]{seber2012linear} using
\begin{equation}
  \label{FindLambdas:residuals}
  \operatorname{e}_{(i)} = \frac{y_i-\hat{y_i}}{1-h_{ii}}\ ,
\end{equation}
where \(y_i\) is the \(i\)th element of the vector \(\bm{y}\), \(\hat{\bm y}_i\) is the \(i\)th element of the vector \(\hat{\bm y} = \bm{H}\bm{y}\), and \(h_{ii}\) is the \(i\)th diagonal element of the hat matrix \(\bm{H} = \bm{X}(\bm{X}'\bm{X})^{-1}\bm{X}'\).
Therefore, we can use the well-known formula for cross-validation in linear regression \citep[see, for example,][p.45]{RWC02}:
\begin{equation}
  \label{FindLambdas:CV}
  \text{CV} = \sum_{i=1}^{n}\left(\frac{y_i-\hat{y_i}}{1-h_{ii}}\right)^2.
\end{equation}

STR finds the optimal \(\lambda\) and \(\theta\) smoothing parameters by minimising CV. The problem of minimising CV can be complex, for example if there are many local minima, and we have no method that guarantees the finding of the global minimum. However, rather simple methods often work well in practice. We perform such optimisation using the Nelder-Mead method as implemented in the \texttt{optim()} function from the stats package in R \citep{R}.

@@

\begin{figure}[!b]\FIGURE
{\includegraphics[width=.9\textwidth]{str_ijds_files/figure-latex/supermarket-1.pdf}}
{\label{fig:supermarket}Supermarket and grocery store log revenue in New South Wales, decomposed with STR. The original data (logged) are shown in the top panel, while the components are shown in the other panels, with 95\% confidence intervals for the trend and season components in blue}
{}
\end{figure}

@@

We obtained similar results using other decomposition methods applied to the same data including STL \citep{cleveland1990stl} and X-13-ARIMA-SEATS \citep{findley2005some}.

%\FloatBarrier

@@

This data set has two seasonal patterns. The first pattern is a weekly seasonal pattern (\(m_1=48\times7=336\)) that represents the specific demand features that are attributable to a particular day of the week. The second pattern is a daily seasonal pattern (\(m_2=48\)) with the topology in Figure \ref{fig:topology_pdf}, which allows the model to distinguish between working days and holidays/weekends, and to make transitions between them. The pattern reflects the tendency to have a higher electricity demand during standard working hours and a lower demand at night. It also reflects the tendency to have different demand patterns on working days and holidays/weekends. A longer series would also have an annual seasonal pattern, but with only 115 days, this cannot be distinguished from the trend component.

\begin{figure}\FIGURE
{\includegraphics[width=.9\textwidth]{str_ijds_files/figure-latex/vic-1.pdf}}
{\label{fig:vic}Electricity demand in Victoria, decomposed using STR. The original data are plotted in black in the first panel, with the decomposed trend plotted in red. The subsequent panels contain the weekly seasonal pattern; the daily seasonal pattern with complex topology; the effect of temperature in Melbourne, which has a daily seasonal pattern with complex topology; the effect of squared temperatures in Melbourne, which is allowed to vary over time; and the residuals. Grey vertical lines represent the residuals with the largest absolute values.}
{}
\end{figure}

@@

\begin{table}\TABLE
{\label{tab:outlierstable}The ten residuals that are largest in absolute value after an STR decomposition.}
{\begin{tabular}[t]{rlcr}
\toprule
Date & Day of week & Time period & Residual\\
\midrule
3 February 2000 & Thursday & 14:30 -- 15:00 & 687.3\\
3 February 2000 & Thursday & 15:00 -- 15:30 & 871.3\\
3 February 2000 & Thursday & 15:30 -- 16:00 & 937.0\\
3 February 2000 & Thursday & 16:00 -- 16:30 & 767.2\\
3 February 2000 & Thursday & 16:30 -- 17:00 & 580.4\\

4 February 2000 & Friday & 16:00 -- 16:30 & -569.7\\
10 February 2000 & Thursday & 14:30 -- 15:00 & -611.4\\
10 February 2000 & Thursday & 15:00 -- 15:30 & -658.9\\
10 February 2000 & Thursday & 15:30 -- 16:00 & -695.7\\
10 February 2000 & Thursday & 16:00 -- 16:30 & -728.5\\
\bottomrule
\end{tabular}}
{}
\end{table}

\begin{figure}\FIGURE
{\includegraphics[width=\textwidth]{str_ijds_files/figure-latex/MelTemp-1.pdf}}
{{\label{fig:MelTemp}The temperature in Melbourne on 1 February 2000, and the following ten days. Ten red lines (two groups of five lines each) mark the times of the outliers. The first group spans 14:30--17:00 on 3 February, while the second group spans 15:30--18:00 on the following day.}}
{}
\end{figure}

@@

Similarly, seasonal functional components can be obtained, using Fourier terms for example \citep[reminiscent of][]{de2011forecasting}. This approach leads, with no additional effort, to handling seasonality with a fractional or varying period.

@@

This can be written as a quantile regression \citep{koenker2005quantile}:
\begin{equation}
  \label{RSTR:qr}
  \hat{\bm\eta} = \operatorname{arg\,min}\limits_{\bm\eta} \left\|\bm{y}^{+}-\bm{X}\bm{\eta}\right\|_{L_1}\ ,
\end{equation}
where \(\bm{y}^{+}\), \(\bm{X}\) and \(\bm\eta\) are defined as in \eqref{STR:linear_model}. This can be solved numerically using a quantile regression algorithm \citep{quantregR}.
